---
title: 'STA 108 Homework 5: R Part'
author: "Yuan Xie"
date: "2025-05-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,eval = TRUE)
```

## Import Environment and Dataset

```{r}
# import the tidyverse environment
library(tidyverse)
library(MASS)
library(car)
library(rcompanion)
# import the csv files with read_csv()
alcohol = read_csv("alcohol.csv")
salary3 = read_csv("salary3.csv")
# briefly know the structures of the csv file by head()
alcohol %>% head()
salary3 %>% head()
```

---

## I
### (a)
```{r}
# construct the linear model
model_alcohol <- lm(BAC ~ BrAC, data = alcohol)

# try the value of lambda between -2 and 2 every 0.1 unit
boxcox_result <- boxcox(model_alcohol, lambda = seq(-2, 2, 0.1))
# find the value of lambda which maximizes the log-likelihood
best_lambda <- boxcox_result$x[which.max(boxcox_result$y)]
best_lambda

if (best_lambda == 0) {
  alcohol <- alcohol %>% mutate(BAC_trans = log(BAC))
} else {
  alcohol <- alcohol %>% mutate(BAC_trans = (BAC^best_lambda - 1) / best_lambda)
}

# plot the graph
alcohol %>% ggplot(aes(x = BrAC, y = BAC_trans)) +
  geom_point() +
  labs(title = paste("Transformed BAC (lambda =", round(best_lambda, 3), ") vs. BrAC"),
       x = "BrAC (mg/L/hour)",
       y = "Transformed BAC")
```
**Interpretation:** The value of $\lambda$ is -0.6666667. The scatter plot of the transformed Y vs. X is shown above.

### (b)
**Hypotheses:** $H_0:$ The residuals have a constant variance. VS. $H_A:$ The residuals do not have a constant variance.
```{r}
alcohol_divide <- alcohol %>%
  mutate(BrAC_group = ntile(BrAC, 4))  # cut into 4 groups

model_trans <- lm(BAC_trans ~ BrAC, data = alcohol_divide)
# use Brown-Forsythe test
the.BFtest <- leveneTest(residuals(model_trans) ~ factor(BrAC_group), data = alcohol_divide, center = median)
the.BFtest
```
**Interpretation:** The p-value of the Brown-Forsythe is 0.7644, which is much greater than $\alpha = 0.05$. We fail to reject $H_0$. The residuals are considered to have a constant variance.

### (c)
```{r}
# do the tukey transform
alcohol_tukey <- alcohol %>%
  mutate(
    BrAC_tukey = transformTukey(BrAC, plotit = FALSE),
    BAC_tukey  = transformTukey(BAC, plotit = FALSE)
  )

# plot the graph
alcohol_tukey %>% ggplot(aes(x = BrAC_tukey, y = BAC_tukey)) +
  geom_point() +
  labs(title = "Scatter Plot of Tukey-Transformed BrAC vs BAC",
       x = "Tukey Transformed BrAC",
       y = "Tukey Transformed BAC")
```
**Interpretation:** Both X and Y are transformed using the Tukey method. The scatter plot of the transformed X and Y is shown above.

### (d)
**Hypotheses:** $H_0:$ The residuals are normally distributed. VS. $H_A:$ The residuals are not normally distributed.
```{r}
model_tukey <- lm(BAC_tukey ~ BrAC_tukey, data = alcohol_tukey)
shapiro_result <- shapiro.test(residuals(model_tukey))
shapiro_result
```
**Interpretation:** The p-value of the Shapiro-Wilk normality test is 0.008138, which is much smaller than $\alpha = 0.05$. We reject $H_0$ and this means the residuals are not normally distributed.  
  
**Hypotheses:** $H_0:$ The residuals have a constant variance. VS. $H_A:$ The residuals do not have a constant variance.
```{r}
alcohol_tukey <- alcohol_tukey %>%
  mutate(BrAC_group = ntile(BrAC_tukey, 4))

# Brown-Forsythe Test
bf_result <- leveneTest(residuals(model_tukey) ~ factor(BrAC_group), data = alcohol_tukey, center = median)
bf_result
```
**Interpretation:** The p-value of the Brown-Forsythe test is 0.6956, which is much greater than $\alpha = 0.05$. We fail to reject $H_0$. The residuals are considered to have a constant variance.

### (e)
**We have not done the Shapiro-Wilk normality test for the Box-Cox case yet.**  
**Hypotheses:** $H_0:$ The residuals are normally distributed. VS. $H_A:$ The residuals are not normally distributed.
```{r}
model_boxcox <- lm(BAC_trans ~ BrAC, data = alcohol)
shapiro_boxcox <- shapiro.test(residuals(model_boxcox))
shapiro_boxcox$p.value
```
**Interpretation:** The p-value of the Shapiro-Wilk normality test is 0.003336245, which is much smaller than $\alpha = 0.05$. We reject $H_0$ and this means the residuals are also not normally distributed for the Box-Cox case.  
**Answer:** Neither model achieves normally distributed residuals according to the Shapiro–Wilk tests. Tukey transformation seems better because it transforms both variables symmetrically and results in better linearity in the scatter plot.

---

## II
### (a)
**Hypotheses:** $H_0: \beta_{4} = 0$  VS. $H_A:$ At least one $\beta_{4} \ne 0$
```{r}
# full model with rk
model_salary3_full <- lm(sl ~ yd + dg + sx + rk, data = salary3)

# Rrduced model without rk
model_salary3_reduced_a <- lm(sl ~ yd + dg + sx, data = salary3)

# compare models using general linear F-test
anova_result <- anova(model_salary3_reduced_a, model_salary3_full)
anova_result
```
**Interpretation:** $H_0: \beta_{4} = 0$  VS. $H_A:$ At least one $\beta_{4} \ne 0$. The value of $F_s$ is 19.395 and the corresponding p-value is 7.791e-07.

### (b)
**Answer:** The p-value is 7.791e-07, which is much smaller than $\alpha = 0.01$. Hence, we reject $H_0$ and conclude that at least one $\beta_{4} \ne 0$.

### (c)
**Hypotheses:** $H_0: \beta_{2} = \beta_{3} = 0$  VS. $H_A:$ At least one of $\beta_{2}$, $\beta_{3} \ne 0$
```{r}
# full model (still the same)
model_salary3_full <- lm(sl ~ yd + dg + sx + rk, data = salary3)

# reduced model without X2 = dg and X3 = sx
model_salary3_reduced_c <- lm(sl ~ yd + rk, data = salary3)

# general linear F-test
anova_result_c <- anova(model_salary3_reduced_c, model_salary3_full)
anova_result_c
```
**Interpretation:** $H_0: \beta_{2} = \beta_{3} = 0$  VS. $H_A:$ At least one of $\beta_{2}$, $\beta_{3} \ne 0$. The value of $F_s$ is 0.6869 and the corresponding p-value is 0.5082.

### (d)
**Answer:** The p-value is 0.5082, which is much bigger than $\alpha = 0.01$. Hence, we fail to reject $H_0$ and conclude that there is insufficient evidence to suggest that either $\beta_{2}$ or $\beta_{3}$ differs from zero.

### (e)
**Answer:** Based on my observations from (b) and (d), $$Y = \beta_0 + \beta_1 X_1 + \beta_{4,1} X_{4,1} + \beta_{4,2} X_{4,2} + \epsilon$$

---

## III
### (a)
```{r}
# Reduced model: only X1 (yd)
model_x1 <- lm(sl ~ yd, data = salary3)
# Full model: X1 and X4 (yd and rk)
model_x1_x4 <- lm(sl ~ yd + rk, data = salary3)

SSE_x1 <- sum(residuals(model_x1)^2)
SSE_x1_x4 <- sum(residuals(model_x1_x4)^2)
reduction <- (SSE_x1 - SSE_x1_x4)/SSE_x1
reduction
```
**Interpretation:** The additional reduction in error we expect to see is 0.5724403.

### (b)
```{r}
model_x4 <- lm(sl ~ rk, data = salary3)

# calculate SSE with only x4
SSE_x4 <- sum(residuals(model_x4)^2)

# partial R^2 of X1 given X4
R2_partial <- (SSE_x4 - SSE_x1_x4) / SSE_x4
R2_partial
```
**Interpretation:** $R^2(X_1 \mid X_4) = 0.05276751$. It means that in addition to the variation explained by "rk", the variable "yd" (years since degree) explains an additional 5.28% of the variation in salary.

### (c)
```{r}
# reduced model with only X1 and X4
model_x1_x4 <- lm(sl ~ yd + rk, data = salary3)
# Full model: all the four predictors
model_x1_x2_x3_x4 <- lm(sl ~ yd + dg + sx + rk, data = salary3)

# calculate SSE
SSE_x1_x4 <- sum(residuals(model_x1_x4)^2)
SSE_full <- sum(residuals(model_x1_x2_x3_x4)^2)

# Reduction in SSE
reduction_c <- (SSE_x1_x4 - SSE_full)/SSE_x1_x4
reduction_c
```
**Interpretation:** The additional reduction in error we expect to see is 0.02900112.

### (d)
```{r}
model_x2_x3 <- lm(sl ~ dg + sx, data = salary3)

# full model SSE has been calculated before
# SSE_x2_x3
SSE_x2_x3 <- sum(residuals(model_x2_x3)^2)

# partial R^2
R2_partial_d <- (SSE_x2_x3 - SSE_full) / SSE_x2_x3
R2_partial_d
```
**Interpretation:** $R^2(X_1, X_4 \mid X_2, X_3) = 0.7578261$. It means that in addition to the variation explained by "dg" and "sx", the variables "yd" and "rk" explain an additional 75.78% of the variation in salary.

### (e)
**Answer:** The results from Problem III agree with the best model selected in II(e). We found that "yd" and "rk" explain an additional 75.78% of the variation in salary. The two predictors are far beyond the explanatory power of the other two. This is consistent with the model $Y = \beta_0 + \beta_1 X_1 + \beta_{4,1} X_{4,1} + \beta_{4,2} X_{4,2} + \epsilon$ where only "yd" and "rk" are included as predictors.

---

## IV
### (a)
```{r}
model.formulas <- c(
  "sl ~ yd", 
  "sl ~ dg", 
  "sl ~ sx", 
  "sl ~ rk", 
  "sl ~ yd + dg", 
  "sl ~ yd + sx", 
  "sl ~ yd + rk", 
  "sl ~ yd + dg + rk",
  "sl ~ yd + sx + rk", 
  "sl ~ yd + dg + sx", 
  "sl ~ yd + dg + sx + rk"
)

PRESS <- function(model) {
  e <- resid(model)
  h <- hatvalues(model)
  sum((e / (1 - h))^2)
}

# All.Criteria from the handout
All.Criteria <- function(the.model) {
  p = length(the.model$coefficients)
  n = length(the.model$residuals)
  the.LL = logLik(the.model)
  the.BIC = -2 * the.LL + log(n) * p
  the.AIC = -2 * the.LL + 2 * p
  the.PRESS = PRESS(the.model)
  the.R2adj = summary(the.model)$adj.r.squared
  the.results = c(LL = the.LL, p = p, n = n, AIC = the.AIC, BIC = the.BIC,
                  PRESS = the.PRESS, R2adj = the.R2adj)
  return(the.results)
}

# apply All.Criteria to all models
criteria.results <- lapply(model.formulas, function(f) {
  model <- lm(as.formula(f), data = salary3)
  All.Criteria(model)
})

# combine results into a data frame
criteria.df <- as.data.frame(do.call(rbind, criteria.results))
rownames(criteria.df) <- model.formulas

# display results
criteria.df
```
**Interpretation:** The model fit critera are listed above.

### (b)
```{r}
# the smaller, the better
criteria.df %>% arrange(BIC) %>% dplyr::select(BIC)
```
**Answer:** According to BIC, sl ~ rk is the best model.

### (c)
```{r}
# the smaller, the better
criteria.df %>% arrange(AIC) %>% dplyr::select(AIC)
```
**Answer:** According to AIC, sl ~ yd + rk is the best model.

### (d)
```{r}
# the smaller, the better
criteria.df %>% arrange(PRESS) %>% dplyr::select(PRESS)
```
**Answer:** According to PRESS, sl ~ yd + rk is the best model.

### (e)
```{r}
# the bigger, the better
criteria.df %>% arrange(desc(R2adj)) %>% dplyr::select(R2adj)
```
**Answer:** According to $R^2_{\text{adj}}$, sl ~ yd + sx + rk is the best model.

### (f)
**Answer:** The models sl ~ yd + rk and sl ~ yd + sx + rk are chosen. PRESS and $R^2_{\text{adj}}$ are used for "prediction" models. sl ~ yd + rk has the smallest PRESS while sl ~ yd + sx + rk has the largest $R^2_{\text{adj}}$. They balances goodness of fit and model complexity well. They reduce the risk of overfitting by excluding unnecessary predictors.

### (g)
**Answer:** The model sl ~ rk is chosen because it has the smallest BIC. BIC tends to pick "correct" models.











